---
title: "Work-life and work-work balance in universities and research centers in Russia. What can we do to stop reinforcing gender gaps?"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,  warning = FALSE, message =  FALSE)
# knitr::opts_chunk$set(fig.pos = 'p') # Places figures on pages separate from text
# knitr::opts_chunk$set(out.width = '70%', dpi=300) # Figure resolution and size
# knitr::opts_chunk$set(fig.env="figure") # Latex figure environment
```

```{r packages}
library(tidyverse)
library(haven)
library(readr)
library(table1)
library(sjlabelled)
library(ggpubr)
library(ggridges)
library(ggsignif)
library(patchwork)
library(expss)
library(FactoMineR)
library(factoextra)
library(RColorBrewer)
library(stringr)
library(stringi)
library(tibble)
library(ggpmisc)
library(readxl)
library(DT)
library(wesanderson)
library(likert)
library(ggrepel)
library(kableExtra)
#library(glmmTMB)
#library(sjPlot)
#library(sjmisc)
#library(jtools)
#library(ggstance)
#library(huxtable)
#library(MASS)
#library(pscl) 
```

```{r functions}
pvalue <- function(x, ...) {
    # Construct vectors of data y, and groups (strata) g
    y <- unlist(x)
    g <- factor(rep(1:length(x), times=sapply(x, length)))
    if (is.numeric(y)) {
        # For numeric variables, perform a standard 2-sample t-test
        p <- t.test(y ~ g)$p.value
    } else {
        # For categorical variables, perform a chi-squared test of independence
        p <- chisq.test(table(y, g),correct = TRUE, simulate.p.value = TRUE, B = 10000)$p.value
    }
    # Format the p-value, using an HTML entity for the less-than sign.
    # The initial empty string places the output on the line below the variable label.
    c("", sub("<", "&lt;", format.pval(p, digits=3, eps=0.001)))
}

all_test <- function(x, ...) {
    # Construct vectors of data y, and groups (strata) g
    y <- unlist(x)
    g <- factor(rep(1:length(x), times=sapply(x, length)))
    if (is.numeric(y)) {
        # For numeric variables, perform a standard 2-sample t-test
        p <- t.test(y ~ g)$p.value
        t <- t.test(y ~ g)$statistic
        par <- t.test(y ~ g)$parameter
    } else {
        # For categorical variables, perform a chi-squared test of independence
        p <- chisq.test(table(y, g))$p.value
        t <- chisq.test(table(y, g))$statistic
        par <- chisq.test(table(y, g))$parameter
    }
    # Format the p-value, using an HTML entity for the less-than sign.
    # The initial empty string places the output on the line below the variable label.
    c("", paste0(sub("<", "&lt;", format.pval(p, digits=3, eps=0.001)), "; val test stat = " , round(t, digits=3), "; df =" , round(par, digits=3)))
}
```

```{r data}
# data wrap in new_search.rmd
set <- read_csv("data/set.csv")

df <- set %>% 
  filter(wave == 'wave_1_2'| wave == 'wave_2_5' | wave ==  'wave_2_6') %>% 
  filter(age_int >= 21 & age_int <= 80) %>% 
  filter(type_doj != "Другое")
```

#### Elena Chechik {.unnumbered}

*Center for Institutional Analysis of Science & Education European University at St. Petersburg*

<br />
<br />


###### Figure S1: Visualization of the leak of men from academia in the 90s and early 00s: Two-humped distribution by age for male economists with the drop in age 50-55. 

```{r}
#| fig-align: 'center'

p <- df %>% 
  ggplot(aes(age_int, group = sex, fill = sex)) +
  geom_density(alpha = 0.8, position = 'identity') +
  scale_fill_manual(values = c("#ffbf49", "#a8a8a8"), guide = FALSE) +
  labs(fill="", x = "", y = "Density") +
  theme_test() + 
  theme(legend.position = 'none',
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.text.x = element_text(size=20),
        axis.title.y = element_text(size=15))

data.tb <- tibble(x = 80, y = 70, 
                  plot = list(p + theme_test(8) + 
                                theme(legend.position = 'none',
                                      plot.background = element_rect(fill = "NA"),
                                      axis.text.y = element_blank(),
                                      axis.ticks.y = element_blank(),
                                      axis.text.x = element_text(size=11),
                                      axis.title.y = element_text(size=11))))

df %>% 
  ggplot(aes(age_int, group = sex, fill = sex)) +
  geom_plot(data = data.tb, aes(x, y, label = plot)) +
  geom_bar(position = position_dodge(preserve = "single")) +
  scale_fill_manual(values = c("#ffbf49", "#a8a8a8")) +
  labs(fill="", x = "Age of respondent", y = "Number of respondents") +
  theme_test() + 
  theme(legend.position = "top",
        axis.text.x = element_text(size = 8),
        panel.grid.major.y = element_line(size = 0.1, 
                                          color = "grey35", 
                                          linetype = "solid"))
```

<br />

###### Figure S2: Percentage of teachers in tertiary education who are female (%)

```{r data2}
NATMON_DS_18112022080205161 <- read_csv("data/NATMON_DS_18112022080205161.csv")
```

```{r}
#| fig-align: 'center'

cont <- c('Germany','Kazakhstan','Finland', 'United States of America','Switzerland', 'Spain', 'Russian Federation')

temp.dat <- NATMON_DS_18112022080205161 %>% filter(Country %in% cont) %>% 
  filter(Time >= 2000) 

ggplot() + 
  geom_hline(yintercept = 50, color = "red", linetype = "dashed", alpha = 0.8) +
  geom_line(data = temp.dat, aes(Time, Value, color = LOCATION)) +
  geom_point(data = temp.dat, aes(Time, Value, color = LOCATION)) +
  geom_text(data = temp.dat %>% filter(Time == last(Time)), 
            aes(label = LOCATION, 
                x = Time + 0.9, 
                y = Value, 
                color = LOCATION)) + 
  guides(color = FALSE) +
  scale_x_continuous(breaks = scales::pretty_breaks(10)) +
  scale_color_brewer(palette = "Dark2", direction = -1) +
  labs(x = "" , y = "Share of women (%)",
       caption = "Data: UNESCO Institute for Statistics (UIS)") +  
  ylim(15, 85) +
  theme_test() + 
  theme(axis.title.x = element_blank(),
        panel.grid.major.y = element_line(size = 0.1, color = "grey70", linetype = "solid"),
        legend.position = "right",
        legend.title = element_blank(),
        plot.caption = element_text(face = "italic"))
```

<br />

###### Table S1: Distribution of women and men by age of having first child and PhD/DS defense

```{r}
Tab_1 <- table1(~ age_int  + 
         first_ch_age + 
         kan_age + doc_age | sex,
  data = df,
  overall = FALSE,
  extra.col = list(`P-value` = pvalue))

Tab_1 <- as.data.frame(Tab_1)[-1, ]
Tab_1[1, 1] <-  "Age"
Tab_1[4, 1] <-  "Age of having a first child"
Tab_1[8, 1] <-  "Age of a PhD defense"
Tab_1[12, 1] <-  "Age of a DS defense"

Tab_1[2, 4] <-  "< 0.001"
Tab_1[5, 4] <-  "< 0.001"
Tab_1[9, 4] <-  "0.224"
Tab_1[13, 4] <-  "0.377"
rownames(Tab_1) <- NULL

kable(Tab_1,
      align = c("l","c", "c", "r"),
      col.names = c(" ","Female<br/> (N=1494)",
                    "Male<br/> (N=901)",
                    "p-value"),
      escape = FALSE) %>%
  kable_styling(full_width = F, position = "center", font_size = 13)
```

<br />

##### Figure S3: Dynamics of age at first birth for women

```{r}
#| fig-align: 'center'

ros <- read_delim("data/rosstat.csv", 
                  delim = ";", escape_double = FALSE, trim_ws = TRUE)
ros <- ros %>% 
  pivot_longer(c(`19 и <`:`не указан \nnot stated`), names_to = "age", values_to = "count")

ros1 <- ros %>% mutate(sh = 100 * count / `\nTotal`,
               sh = round(sh, digits = 2)) %>% 
  filter(`\nYears` >= 2000) %>% 
  mutate(age = if_else(age == "50 и >", "50 и >", age)) %>% 
  filter(age != "не указан \nnot stated") 

ggplot() +
  geom_line(data = ros1, aes(`\nYears`, sh, group = age, color = age )) +
  geom_point(data = ros1, aes(`\nYears`, sh, group = age, color = age )) +
  geom_text_repel(data = ros1 %>% filter(`\nYears` == last(`\nYears`)), 
                  aes(label = age, 
                      x = `\nYears` + 1,
                      y = sh,
                      color = age), 
                  nudge_x = 0, direction = "y", 
                  hjust = "center", box.padding = -0.01) +
  guides(color = FALSE) +
  scale_x_continuous(breaks = scales::pretty_breaks(10)) +
  scale_color_brewer(palette = "Dark2", direction = -1) +
  labs(x = "" , y = "Share of age group (%)",
       caption = "Data: Russian Statistical Agency") +
  theme_test() + 
  theme(axis.title.x = element_blank(),
        panel.grid.major.y = element_line(size = 0.1, color = "grey70", linetype = "solid"),
        legend.position = "right",
        legend.title = element_blank(),
        plot.caption = element_text(face = "italic")) 
```
